Working title, The impact of heterospecific pollen on plant reproductive success is mediated by phylogenetic distance and floral reproductive traits

### 2018-12-03 Manuscript_draft/Manuscript_draft.Rmd
Writing day

In natural systems plant species normally coexist and share their floral visitors with other species @bascompte2003. This pollinator sharing from the plant perspective at the pre-pollination stage can be negative due to competition @pauw2013 or positive due to facilitation @carvalheiro2014. Once the floral visitor has arrived to the flower, pollen deposition on the stigma can take place and hence ovule fertilization. An increasing number of visits generally correlates with higher chances of fertilization @engel2003. However this is not always the case, among these possible flower visitors we find also nectar robbers and pollen thiefs @inouye1980 and the quality of pollen that is deposited on the stigma is also highly relevant to the pollination succes @aizen2007. Moreover, other less study issues in the pollination process are conspecific pollen loss and the arrival of foreign pollen which can have important detrimental effects for the species fitness @morales2008 @ashman2013.

@@

Recent studies have advanced in the ecological importance of heterospecific pollen effect @morales2008 @ashamn2013 @arceo2016. A general overview of heterospecific pollen effect is that it can play an important role on species fitness but seems to be context dependent and not always fitness decreased is produced @morales2008. There are several traits that are postulated to be highly relevant to this process but the correlation with heterospecific pollen effect remains porrly understood.

Mention relatedness somwhere here.

Moreover, the different methods used make difficult to compare among different experiments. For instance many articles do not describe the foreign pollen quantity used to test for effect or how this pollen was applied (Refs). The small size of pollen, great environmental variability and population outcrossing rates among others make of this field of study a daunting task and higly unpredictable.

>>>>

In natural systems plant species normally coexist and share their floral visitors with other species @bascompte2003. This pollinator sharing from the plant perspective at the pre-pollination stage can be negative due to competition @pauw2013 or positive due to facilitation @carvalheiro2014. Once the floral visitor has arrived to the flower, pollen deposition on the stigma can take place and hence ovule fertilization. An increasing number of visits generally correlates with higher chances of fertilization @engel2003. However this is not always the case, among these possible flower visitors we find also nectar robbers and pollen thiefs @inouye1980 and the quality of pollen that is deposit on the stigma is also highly relevant to the pollination succes @aizen2007. Moreover, other less study issues in the pollination process are conspecific pollen loss and the arrival of foreign pollen which can have important detrimental effects on species fitness @morales2008 @ashman2013.

@@

Recent studies have advanced in the ecological importance of heterospecific pollen effect @morales2008 @ashman2013 @arceo2016. A general overview of foreign pollen arrival is that it can play an important role on species fitness but seems to be context dependent and not always produce a decrease in fitness @morales2008. Part of this unpredictability is due to the enormous variability of foreing pollen transferred in nature, where levels between 0 and 75 percent are seen, but most commonly values ranges between 10 and 20 percent of the total pollen load, being the generalist species the ones that receive greater loads of heterospecific pollen @montgomery2012 @fang2013. Although heterospecific pollen quantity is fundamental to understand the outcome of the interaction so is the different traits of both pollen donor and recipient. @ashman2013 postulated the first predictive framework for traits of heterospecific pollen effect where traits such as compatibility system and pollen size among others should be crucial to understand foreing pollen effect. Moreover,  in @tong2016 a assymetric effect was shown in a crossing experiment between 6 species of the genus *Pedicularis* where the pollen of long styled species was able to grow the ful length of the syle on short styled species but not viceversa. Despite these recent caveats, we still lack empirical evidence to affirm what are the main traits that drive heterospecific pollen effect for both pollen donor and recipient at seed production level. Interestingly, this trait based question cannot be solved without consider the phylogenetic relatedness of the species. A commonly understanding is that close related species will have greater negative impact

### 2018-12-03 Manuscript_draft/Manuscript_draft.tex
Writing day

In natural systems plant species normally coexist and share their floral
visitors with other species Bascompte et al. (2003). This pollinator
sharing from the plant perspective at the pre-pollination stage can be
negative due to competition Pauw (2013) or positive due to facilitation
Carvalheiro et al. (2014). Once the floral visitor has arrived to the
flower, pollen deposition on the stigma can take place and hence ovule
fertilization. An increasing number of visits generally correlates with
higher chances of fertilization Engel and Irwin (2003). However this is
not always the case, among these possible flower visitors we find also
nectar robbers and pollen thiefs Inouye (1980) and the quality of pollen
that is deposited on the stigma is also highly relevant to the
pollination succes Aizen and Harder (2007). Moreover, other less study
issues in the pollination process are conspecific pollen loss and the
arrival of foreign pollen which can have important detrimental effects
for the species fitness Morales and Traveset (2008) Ashman and
Arceo-Gómez (2013).

@@

Recent studies have advanced in the ecological importance of
heterospecific pollen effect Morales and Traveset (2008)
({\textbf{???}}) Arceo-Gómez and Ashman (2016). A general overview of
heterospecific pollen effect is that it can play an important role on
species fitness but seems to be context dependent and not always fitness
decreased is produced Morales and Traveset (2008). There are several
traits that are postulated to be highly relevant to this process but the
correlation with heterospecific pollen effect remains to be tested.
Moreover, the different methods used make difficult to compare among
different experiments. For instance many articles do not describe the
foreign pollen quantity used to test for effect or how this pollen was
applied (Refs). The small size of pollen, great environmental
variability and population outcrossing rates among others make of this
field of study a daunting task and higly unpredictable.

@@

@@

@@

>>>>

In natural systems plant species normally coexist and share their floral
visitors with other species Bascompte et al. (2003). This pollinator
sharing from the plant perspective at the pre-pollination stage can be
negative due to competition Pauw (2013) or positive due to facilitation
Carvalheiro et al. (2014). Once the floral visitor has arrived to the
flower, pollen deposition on the stigma can take place and hence ovule
fertilization. An increasing number of visits generally correlates with
higher chances of fertilization Engel and Irwin (2003). However this is
not always the case, among these possible flower visitors we find also
nectar robbers and pollen thiefs Inouye (1980) and the quality of pollen
that is deposit on the stigma is also highly relevant to the pollination
succes Aizen and Harder (2007). Moreover, other less study issues in the
pollination process are conspecific pollen loss and the arrival of
foreign pollen which can have important detrimental effects on species
fitness Morales and Traveset (2008) Ashman and Arceo-Gómez (2013).

@@

Recent studies have advanced in the ecological importance of
heterospecific pollen effect Morales and Traveset (2008) Ashman and
Arceo-Gómez (2013) Arceo-Gómez and Ashman (2016). A general overview of
foreign pollen arrival is that it can play an important role on species
fitness but seems to be context dependent and not always produce a
decrease in fitness Morales and Traveset (2008). Part of this
unpredictability is due to the enormous variability of foreing pollen
transferred in nature, where levels between 0 and 75 percent are seen,
but most commonly values ranges between 10 and 20 percent of the total
pollen load, being the generalist species the ones that receive greater
loads of heterospecific pollen Montgomery and Rathcke (2012) Fang and
Huang (2013). Although heterospecific pollen quantity is fundamental to
understand the outcome of the interaction so is the different traits of
both pollen donor and recipient. Ashman and Arceo-Gómez (2013)
postulated the first predictive framework for traits of heterospecific
pollen effect where traits such as compatibility system and pollen size
among others should be crucial to understand foreing pollen effect.
Moreover, in Tong and Huang (2016) a assymetric effect was shown in a
crossing experiment between 6 species of the genus \emph{Pedicularis}
where the pollen of long styled species was able to grow the ful length
of the syle on short styled species but not viceversa. Despite these
recent caveats, we still lack empirical evidence to affirm what are the
main traits that drive heterospecific pollen effect for both pollen
donor and recipient at seed production level. Interestingly, this trait
based question cannot be solved without consider the phylogenetic
relatedness of the species. A commonly understanding is that close
related species will have greater negative impact

@@

\hypertarget{ref-fang2013}{}
Fang, Q., and S.-Q. Huang. 2013. A directed network analysis of
heterospecific pollen transfer in a biodiverse community. Ecology
94:1176--1185.

@@

\hypertarget{ref-montgomery2012}{}
Montgomery, B. R., and B. J. Rathcke. 2012. Effects of floral
restrictiveness and stigma size on heterospecific pollen receipt in a
prairie community. Oecologia 168:449--458.

@@

\hypertarget{ref-tong2016}{}
Tong, Z.-Y., and S.-Q. Huang. 2016. Pre-and post-pollination interaction
between six co-flowering pedicularis species via heterospecific pollen
transfer. New Phytologist 211:1452--1461.
